Data Cleaning

### 2025-12-02 Data Cleaning.Rmd
Added code for cleaning temperature, pH, conductivity, and turbidity data.

>>>>

```{r}
# output merged data to CSV (run after processing)

library(readr)
write_csv(merged_clean2, "alkalinity_hardness_cleaned.csv")

## Raw Water Data

```{r}

library(lubridate)
library(ggplot2)
library(zoo)
library(purrr)

## Import data, convert timestamp
rwar <- read.csv("RWAR_turb_cond_counts_pH_temp.csv")
rwar2 <- rwar |> mutate(DateTime = mdy_hm(rwar$Timestamp))
summary(rwar2)

## Temperature Data
# Plot
ggplot(rwar2) + geom_point(aes(x=DateTime, y=Temperature..deg.C))
# Find region of approx. zero temperature (outliers), set to NA
# Interpolate between nearest non-NA temperatures where temp is NA
rwar2_temp <- rwar2 |>
  select(c("DateTime", Temperature..deg.C)) |>
  mutate(temp_cleaned = ifelse(Temperature..deg.C < 1, NA, Temperature..deg.C)) |>
  mutate(temp_interp = na.approx(temp_cleaned, na.rm = FALSE))
# Average by day
rwar2_temp_daily <- rwar2_temp |>
  mutate(Day=date(DateTime)) |>
  group_by(Day) |>
  summarise(mean_Temp = mean(temp_interp))
ggplot(rwar2_temp_daily) + geom_point(aes(x=Day, y=mean_Temp))

## pH Data
# Plot
ggplot(rwar2) + geom_point(aes(x=DateTime, y=pH..SU))
# Does not seem like there are huge outliers, so average by day
rwar2_pH_daily <- rwar2 |>
  select(c("DateTime", pH..SU)) |>
  mutate(Day=date(DateTime)) |>
  group_by(Day) |>
  summarise(mean_pH = mean(pH..SU))
ggplot(rwar2_pH_daily) + geom_point(aes(x=Day, y=mean_pH))

## Conductivity Data
# Plot
ggplot(rwar2) + geom_point(aes(x=DateTime, y=Conductivity..mS.cm.1))
# Remove obvious outliers (only a few measurements; does not remove full days)
rwar2_cond <- rwar2 |> 
  select(c("DateTime", Conductivity..mS.cm.1)) |>
  filter(Conductivity..mS.cm.1 < 2 & Conductivity..mS.cm.1 > 0.2)
ggplot(rwar2_cond) + geom_point(aes(x=DateTime, y=Conductivity..mS.cm.1))
# Average by Day
rwar2_cond_daily <- rwar2_cond |>
  mutate(Day=date(DateTime)) |>
  group_by(Day) |>
  summarise(mean_Conductivity = mean(Conductivity..mS.cm.1))
ggplot(rwar2_cond_daily) + geom_point(aes(x=Day, y=mean_Conductivity))

## Turbidity Data
# Plot
ggplot(rwar2) + geom_point(aes(x=DateTime, y=Turbidity..NTU))
# Remove obvious outliers (only a few measurements; does not remove full days)
# Main obvious outliers are in mid 2019 and 2020, and are all above 7 NTU
rwar2_turb <- rwar2 |> 
  select(c("DateTime", Turbidity..NTU)) |>
  filter(Turbidity..NTU < 7)
ggplot(rwar2_turb) + geom_point(aes(x=DateTime, y=Turbidity..NTU))
# Average by Day
rwar2_turb_daily <- rwar2_turb |>
  mutate(Day=date(DateTime)) |>
  group_by(Day) |>
  summarise(mean_Turbidity = mean(Turbidity..NTU))
ggplot(rwar2_turb_daily) + geom_point(aes(x=Day, y=mean_Turbidity))

## Merge daily temp, pH, conductivity, turbidity

list_daily_data <- list(rwar2_temp_daily,
                        rwar2_pH_daily, 
                        rwar2_cond_daily,
                        rwar2_turb_daily)
merged_daily_data <- list_daily_data |>
  reduce(full_join, by = "Day") |>
  rename(Timestamp = Day)

```{r}
## Export CSV of cleaned raw water data

write_csv(merged_daily_data, )
```
